Maenpuen et al. Appendix S1-7: Sources and consequences of mismatch between leaf disc and whole-leaf leaf mass per area (LMA)

### 2022-05-13 ms/response_letter.Rmd

---
geometry: margin=1in
link-citations: yes
#csl: templates/xxxx.csl
#bibliography: xxxx.bib
fontsize: 12pt
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \usepackage{lineno}
 # - \linenumbers
output:
  pdf_document:
    keep_md: false
    fig_caption: yes
    keep_tex: no
    toc: no
    number_sections: no
    template: templates/eisvogel2.tex
    #template: null
    pandoc_args:
      - "--filter"
      - "pandoc-crossref"
  word_document:
    fig_width: 6
    fig_caption: yes
    toc: FALSE
    reference_docx: templates/rmd_style.docx
    pandoc_args:
    - "--filter"
    - "pandoc-crossref"
    - "--mathml"
---

Dear Editors,

We hereby resubmit the manuscript (AJB-D-21-00391), “Sources and consequences of mismatch between leaf disc and whole-leaf leaf mass per area (LMA)” (Phisamai Maenpuen, Masatoshi Katabuchi, Yusuke Onoda, Cong Zhou, Jiao-Lin Zhang, Ya-Jun Chen).
We thank the associate editor and the reviewers for their careful attention to detail in our manuscript and for providing comments that have greatly improved the study.

Based on the comments, we have updated the manuscript to reflect the following:

1. To quantify the effect of leaf morphology (leaf tissue density, leaf area, leaf thickness) and leaf punch size on the relationship between whole-leaf and leaf disc LMA, we have built a hierarchical Bayesian model.
This allows to quantify when the whole-leaf LMA to leaf disc LMA ratio become large and in which direction the relationship is changing.

2. We also removed the analysis that grouped species into four categories based on medians of leaf area and leaf thickness because it is somewhat arbitral as pointed by the reviewer.
Instead, the above Bayesian model can was used to test our hypotheses.

3. We have removed the individual-based analysis from the main text because variance among species are not independent in the individual-level data, and thus SMA or OLS regressions are not appropriate.
Analyses of individual-level data requires mixed linear models with species-level errors, but this is not useful for the context of our paper.
We therefore only report species-level results.

The programming code and the data supporting the findings of this study will be deposited at Zenodo (https://zenodo.org) and Github (https://github.com/mattocci27/leaf-disc).
At the moment, we have attached the stan code for our Bayesian model.

Below we provide detailed responses.
The editing history is recorded in a separate pdf file.

We appreciate the opportunity to revise and improve our manuscript, and we thank the editors and reviewers for their time and consideration.

Sincerely,

Masatoshi Katabuchi

Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences

(on behalf of coauthors: Phisamai Maenpuen,Yusuke Onoda, Cong Zhou, Jiao-Lin Zhang, and Ya-Jun Chen)

\newpage

@@

> AE comments: Thanks for the submission. I agree with the reviewers that the paper has merit and will be of interest to many researchers because LMA is a widely used trait. Both reviewers have some feedback for you to consider. Reviewer 1's suggestions would require extensive revisions to address. Reviewer 2 has suggested a modification to the title. I think the existing title is fine if you prefer to keep that but would ask you to carefully consider all other comments carefully.

We appreciate the positive comments, and thank you for the opportunity to revise the manuscript.
We have carefully detailed our responses to each point below.

@@

> Reviewer #1: Summary of Manuscript:
> This manuscript compares two methods of measuring LMA (leaf mass per area), an ecologically important functional trait in plants. It is a carefully executed and useful piece of work, in that LMA is a very widely measured plant trait and many readers will be interested in identifying the best methods to use when measuring LMA in their system, as well as how to compare their own measurements to other published estimates. However, I believe the paper needs modification; I include recommendations, including three major concerns, below:

We thank the reviewer for the encouraging comments and helpful critiques.

> 1. Framing: I think the paper misses an opportunity to provide a bit more context for why the study is important. The authors could provide more concrete examples of why standardizing LMA techniques across studies matters: e.g. what is the ecological significance of an LMA measurement difference between the two methods?  This could also be set up in the introduction by providing more context for LMA as a functional trait. For example, by citing the TRY database, the authors could provide readers with the importance and ubiquity of LMA relative to other plant traits (e.g. of all the studies of plant traits in TRY, how many report LMA?)

Thank you.
We now note that a key difference between leaf disc and whole-leaf LMA is that leaf disc LMA focuses on lamina rather than whole-leaf and usages depend on ecological contexts (Line 75-77).
We also have mentioned the TRY database to explain LMA is the most often used leaf trait in TRY (Line 59-63).

@@

> 2. A crucial thing for readers to know is how to "convert" one method's estimate of LMA to another: the authors could report an equation (or a few, accounting for some additional covariates such as leaf area and leaf thickness, which this paper identifies as important modifiers of the relationship) that scholars could use in the future if trying to compare LMA estimates collected using different methods. In other words, taking the eq from Appendix 2 and modifying it given the results of this study.

Thank you.
This is a good point and an important thing to clarify.
We now use a linear model that includes leaf area and leaf thickness as covariates but we suggest OLS models without covariates for the conversion of LMA because leaf thickness and leaf area affect not only the mean but also the variance in the relationship between the leaf disc and whole-leaf estimates, and that is why R2 values changed among the four groups (thin vs. think and small vs. large leaves) and it is not to practical to convert LMA using leaf area or leaf thickness.
We also present a practical guideline for conversion of LMA.

> 3. I am confused and unconvinced by the inclusion and discussion of intraspecific trait variation on several levels. First, the authors do not clearly explain ITV, its importance, or its relevance to LMA. Secondly, the methods used to quantify ITV are not recommended (see Yang et al 2020. Large underestimation of intraspecific trait variation and its improvements. Frontiers in Plant Science); ideally the authors should use Bao's CV estimator. Finally, there is little interpretation of the results of their analysis. I think this metric needs to either be excluded entirely or significantly expanded upon.

Thank you.
We have cited the recent review by Westerband et al. as well as other works to explain ITV, its importance and the extent of ITV for LMA (Line 104-111).
We now use Bao's CV estimator (CV4) instead of the coefficient of variation of log-normally distributed data ($CV = \sqrt{exp(s_{\mathrm{ln}}^2) - 1}$).
We also limited to the analysis for species that have at least 5 individuals.
The relationship between the two estimates became much weaker (Fig. 4).

@@

> L49 I think the introduction would benefit from an initial sentence to give more context for the definition of LMA (perhaps by restructuring this paragraph to introduce traits and the LES first, and then defining LMA in that context).

Thank you.
We now introduce the LES in the first sentence so that the readers can understand the role of LMA in LES more clearly.

> L69-73 This sentence is crucial for your hypotheses but is buried in the middle of this paragraph. I suggest breaking this paragraph (beginning on line 63) up into 2: the first describing different methods of measurement and how commonly each are used, and any previous attempts to compare them. The second paragraph can get into which leaf types are likely to give similar or different estimates based on the two metrics and why.

Thank you.
We now break up the paragraph into two sections: the first describing the different methods and their record numbers in TRY, and previous works that compared them (Line 64-86), and the second describing the expected effects of leaf morphology on the relationship between whole-leaf and leaf disc LMA (Lin 87-103).

@@

> L154-160 see major comment 3

Thank you.
We now use Bao's CV as noted above.

> L169-172 This is the major result of the paper. I suggest making it the first line of the results section (however I would modify it to fit the individual, rather than species-mean, level data: see below).

Thank you.
We have started the result section with the shift in elevation between whole-leaf and leaf disc LMA (i.e., whole-leaf LMA is greater than leaf disc LMA).
Please see our response for using species- vs. individual-level data below.

@@

> L175-176 Based on this, I am confused about the statistical approach. Did the authors separately run regressions for species-level means and individual level means? If so, it is almost a statistical inevitability that the individual level analysis would have a lower r2 than the species level analysis: therefore I would not report this as a result, per se. In addition, the species level means is more or less meaningless in the context of this paper: much more importantly is the comparison of individual based measurements. I would therefore remove the species level results form the manuscript, and instead only report on the individual level data.

Thank you.
We agree that using both species- and individual-level data in the same statistical framework is confusing and not appropriate.
Instead of using individual-level data, we now only use the species-level data
For individual-level data, variance among species are not independent, and thus SMA or OLS regressions are not appropriate.
Analyses of individual-level data requires mixed linear models with species-level errors, but this is not useful for the context of our paper.
We therefore only report species-level results.

> L182 I believe leaf punch size was something that varied between datasets, and was never measured on the same leaves. I'm therefore confused as to how the authors are interpreting the effect of "leaf punch size" in their models: could this not simply be a reflection of the different species represented in the different datasets?

Thank you.
We now include other continuous variables (leaf tissue density, leaf area, and leaf thickness) as covariates in our model, and thus we have some confidence in the effect of leaf punch.
However, as the reviewer pointed out, we should have used the same leaf punch size or repeated measurements using different leaf punch sizes.
We have discussed the effects of leaf punch size more carefully (Line 273-282).

@@

> L183-184 Which metric had higher ITV? That is a more relevant result than simply the R2 value.

Leaf disc LMA had higher ITV.
We have added the paired t-test for this in the results section (Line 251-255).

@@

> L193 you mention "appropriate calibration": it would be extremely useful if the authors were to report an equation (or several) for converting LMA measured by one metric to the other, ideally that incorporate leaf area and thickness (see major comment).

Thank you.
Now we have the subsection for conversion between the two LMA estimates in the discussion.

@@

> L247 is this reported in the results somewhere? I thought the multiple discs/leaf were measured together?

Thank you.
That is not a result but an explanation.
Since we only measured 3 discs for each leaf in the Yunnan dataset, which is a common protocol, we cannot avoid the effect of variance within leaves.
We have updated the text (Line 364-365).

> Appendix S2 seems to be missing the x axis titles

Thank you.
The two panels share the same x axis, and the units are in the strip.

@@

> Reviewer #2: In this work, the authors compared two methods to measure LMA of plants: leaf disc sampling and whole-leaf sampling using 334 woody species from East Asia. The main finding are that (i) mean values of leaf disc LMA is a good proxy for mean values of whole-leaf LMA; (ii) larger and thinner leaves tended to show the large discrepancies between the leaf disc and whole-leaf estimates. Therefore, the present work implies that when one needs to mix leaf disc LMA and whole-leaf LMA in a single analysis (e.g., meta-analysis), appropriate calibrations are necessary. This manuscript was well written. At least, I can read the whole manuscript fast. The study is valuable for choosing a better approach when calculating LMA. I suggest minor to moderate revisions for this manuscript.

> General Comments:

> In the Introduction section, the authors had better add some introduction to the leaf disc sampling method, since this method was focused to make a comparison with the whole-leaf sampling method for calculating LMA.

Thank you.
We have mentioned several ecological studies that used leaf disc LMA in the introduction (Line 70-77).

@@

> The authors had better explain why they used the LMA values with leaf petioles mass.

Thank you.
We have added the records of TRY database that shows LMA including petioles are very common (Line 64-70).

@@

> Lines 26 and 51: Change "the key trait" to "an important functional trait".

Thank you.
We have updated the first one but keep the second one as "a key trait" because we have added the explanation on why LMA is considered as "a key trait" in the LES at the beginning of the introduction.

> Line 57: I think that the following reference helps to support this statement, because LMA values were found to vary across leaf age, and LMA also exhibits a plasticity to climate when leaves enter the maturation stage (Jiao et al., 2022).
> Jiao, Y., Niklas, K.J., Wang, L., Yu, K., Li, Y., Shi, P. 2022. Influence of leaf age on the scaling relationships of lamina biomass vs. area. Frontiers in Plant Science In press.

Thank you.
We have cited it in the context of "diminishing returns" in the discussion (Line 322-323).

@@

> Lines 97 and 98: Drop the numerical values of R squared. There is no need to exhibit the results here.

Thank you, but to make the statement of "good correlations" more quantitative, we do feel showing *R^2^* values is still useful.
We, therefore, kept the *R^2^* values in the introduction.

> Lines 105 and 106: Why did you only collect the data in China and Japan. Here, I think, you had better add one to three sentences to explain the feasibility of showing general results for other floras.

Thank you.
We have added MAT and MAP in the main text to support that our dataset covers the wide climate ranges in the distribution of broad-leaved tree species (Line 143-147).

@@

> Lines 124 and 125: Why did you maintain leaf petiole in calculating LMA? It is strange for me, because it might inflate LMA given that its definition is dry mass per unit leaf area. What is the meaning for "avoiding thick veins"? Do you mean in the experiment you punch the positions without having the secondary order veins?

Thank you.
The main function of leaves is to deliver a profitable return on the investment that has been made in constructing the leaf (Westoby, Reich and Wright 2013), and thus it is also natural to include leaf petiole in LMA.
As noted above, we have added the records of TRY database that shows LMA including petioles are actually very common.

We now more clearly explain that thick veins are first to secondary order veins (Line 163).

### 2022-05-13 ms/LMA_method.Rmd
add lines

The primary function of leaves is to return photosynthetic revenue on the investment that has been made in constructing the leaf [i.e., leaf economic spectrum, @Wright2004a; @Westoby2013].
Leaf mass per area (LMA or 1/SLA [specific leaf area]), determined by lamina thickness and leaf tissue density (LD), describes how much biomass is invested into given photosynthetic leaf area, which is a key trait in the leaf economic spectrum [@Wright2004a; @Poorter2009; @Onoda2017].
Generally, resource-acquisitive species (fast-growing species) tend to have low LMA values, showing high photosynthesis, high nutrients, and often have fast leaf turnover [@Wright2004a].
In contrast, resource-conservative species (slow-growing species) often have higher LMA values with the opposite patterns [@Garnier1994; @Wright2004a; @Reich2014; @Diaz2016].
The LMA is also frequently used in plant growth analysis [@Evans1972; @Poorter2014; @Falster2018] because relative growth rates (RGR) can be decomposed into the product of net assimilation rate (NAR), leaf mass ratio (LMR) and LMA (i.e., RGR = NAR $\times$ LMR $\times$ LMA^-1^).
Another appealing feature of LMA is that it is relatively easy to measure large numbers of species [i.e., 'soft' trait, @Diaz2004].
Therefore, LMA has been of interests to ecologists and widely used since the first report more than a century ago [@Hanson1917].
Actually, LMA has one of the best species coverage in leaf traits of the TRY plant data database (16,460 species) and was also the most often requested leaf trait (2,977 out of 7,330 requests), followed by leaf by leaf nitrogen (N) contents per leaf dry mass [12,238 species, 1,938 requests, @Kattge2020 [TRY version 5, status October 1 2019]]

@@

In this study, we aimed to investigate the relationship between whole-leaf (including petiole) LMA and leaf disc LMA,
whether the relationship varied with leaf tissue density, leaf area and leaf thickness, and
whether the extent of ITV for each species match between whole-leaf and leaf disc based estimates.
We collected a total of `r val$all_sp` woody species from four biomes (tropics, subtropics, savanna, and warm-temperate) to cover the wide range of geography.
We evaluate the following hypotheses:
(1) whole-leaf LMA would be greater than leaf disc LMA because whole-leaf LMA includes petioles and midribs that have greater dry mass per unit area than laminas [@Niinemets2007],
(2) species with higher leaf tissue density, larger and/or thinner leaves would show larger variance and differences in LMA estimates because whole-leaf LMA should be driven more in those species by veins and petioles that are likely to be ignored in leaf disc samples, and
(3) the extent of ITV does not match well between whole-leaf and leaf disc based estimates because leaf disc samples have more sources of variation than whole-leaf samples.
In addition to leaf morphology, we also investigate the effects of leaf punch size on the relationship between whole-leaf and leaf disc LMA.

@@

There were a significant shift in elevation between whole-leaf and leaf disc LMA (Fig.1, SMA slope: `r val$sma_slope_all_mean` [95%CI: `r val$sma_slope_all_lwr`, `r val$sma_slope_all_upr`], SMA intercept: `r val$sma_int_all_mean` [95%CI: `r val$sma_int_all_lwr`, `r val$sma_int_all_upr`]), suggesting that the whole-leaf estimates of LMA were `r val$sma_per_all_mean`% [95%CI: `r val$sma_per_all_lwr`, `r val$sma_per_all_upr`] greater than disc-based estimates in the arithmetic scale.

@@

Leaf disc LMA explained interspecific variations in whole-leaf LMA (*R^2^* = `r val$r2_lma`, *n* = `r val$all_sp`) quite well, and the strength of their relationship was approximately same or slightly weaker than previously reported  [*R^2^* = 0.92, *n* = 409, for whole-leaves in @Kraft2008; *R^2^* = 0.92, *n* = 364 for leaf laminas in @Onoda2011].
Standardized major axis (SMA) regression analysis showed significantly greater intercepts when all the species were pooled (9.4% differences).
Together, these findings suggest that disc-based estimates can be generally used as good proxies for their whole-leaf estimates with appropriate calibration (see discussion below).
However, leaf tissue density, leaf area, leaf thickness and leaf punch size also influenced the relationship between the whole-leaf LMA and leaf disc LMA (Fig. 2 and Appendix S4).

@@

The effect of leaf punch size was the most important source of the variation (Fig. 2, Appendix S4), and samples obtained with a small leaf punch showed greater variance in whole-leaf to leaf disc LMA ratio.
Small leaf disc size inflates trait varition within leaves because dry mass will be inflated when major veins are unintentionally included in the samples.
Species with lower leaf tissue density and thin leaves showed greater variance in whole-leaf to leaf disc LMA ratio especially when a small punch was used (Fig.2ac: small leaf punch, Fig. 2a: large leaf punch).
@Kraft2008 found high *R^2^* values (*R^2^* = 0.92) between whole-leaf and leaf disc LMA based on leaf discs of approximately 1.0-cm diameter for sapling leaves under closed canopy that are softer and thinner than adult leaves [@Kitajima2010; @Oktavia2020].
Altough a leaf punch of 1.0-cm diameter was used only in the Yakushima dataset in our study, the correlations were also strong for the Yakushima dataset, regardless of leaf tissue density or leaf thickness (Fig. 2ac).
On the other hand, a previous study found that leaf disc LMA obtained with a very small leaf punch (0.3-cm diameter) and whole-leaf LMA had different relationships between shade tolerance [@Poorter2009a], suggesting that leaf disc LMA and whole-leaf LMA can be inherently different when small disc samples were used.
The relatively low *R^2^* values in the Yunnan dataset (Appendix S5) compared to the previous study and the Yakushima dataset may suggest not only the effect of distinct veins on variance in LMA [@John2017] or distinct veins with kite-type appearances [@Grubb1986] but also technical difficulty in measuring tiny weight of discs in lower LMA leaves.
We found that estimates of whole-leaf and leaf disc LMA diverged when total dry mass of leaf discs was small (Appendix S6), and thus species with lower leaf tissue density and thin leaves, which have smaller dry mass (Appendix S7), tended to show grater variation in whole leaf to leaf disc LMA ratio especially when we used a smaller leaf punch.

@@

When the objective is to use leaf disc LMA as an alternative for whole-leaf LMA, using an appropriate size of leaf punch (e.g., 1.0-cm in diameter) is important for obtaining stable estimates of leaf disc LMA that matches well with whole-leaf LMA.
Our results suggest that simple conversion between whole-leaf and leaf disc LMA is difficult for species obtained with a small leaf punch.
The effects of leaf tissue density and leaf area are predictable when we used a large leaf punch (Fig. 2ab: large leaf punch), suggesting that conversion between leaf disc and whole-leaf LMA should be possible using leaf area as a covariate (measuring leaf tissue density requires whole-leaf LMA itself).
OLS regressions with leaf area as a covariate as following:
(1) log~10~LMA~w~ = `r val$yaku2_slope_disc` $\times$ log~10~LMA~d~ + `r val$yaku2_slope_area` $\times$ log~10~LA,
(2) log~10~LMA~d~ = `r val$yaku2_int_inv_log10` + `r val$yaku2_slope_inv_leaf` $\times$ log~10~LMA~w~  - 0.063  $\times$ log~10~LA (*R^2^* = `r val$yaku_r2_area`, *AIC* = `r val$yaku_aic_area`, *n* = `r val$yaku_sp`),
where LMA~w~, LMA~d~, and LA are the whole-leaf LMA, leaf disc LMA and leaf area, respectively.
OLS regressions without leaf area as a covariate as following:
(3) log~10~LMA~w~ = `r val$yaku_int_log10` + `r val$yaku_slope_disc` $\times$ log~10~LMA~d~ or
(4) log~10~LMA~d~ = `r val$yaku_slope_inv_leaf` $\times$ log~10~LMA~w~
(*R^2^* = `r val$yaku_r2_simple`, *AIC* = `r val$yaku_aic_simple`, *n* = `r val$yaku_sp`).
Nonetheless, including leaf area only slightly improves the *R^2^* values for the Yakushima dataset where a large punch was used.
This is because the large discrepancy in the whole-leaf to leaf disc LMA for large leaves can be explained by scaling relationships between leaf mass and leaf area called "diminishing returns" [@Niklas2007] which refers to increases in leaf dry mass including petioles do not result in the proportional increase in leaf area including petioles [@Li2022], which is strong for mature leaves [@Jiao2022].
The "diminishing returns" can be either described as the scaling slope of leaf area in the model 1 or the scaling slope of leaf-disc LMA and the intercept in the model 3.
When one needs to mix leaf disc LMA and whole-leaf LMA in a single analysis (e.g., meta-analysis) or quickly measure leaf disc LMA as a proxy for whole-leaf LMA, we recommend to convert LMA values using one of the models above.

@@

*Conclusions*

>>>>

The primary function of leaves is to return photosynthetic revenue on the investment that has been made in constructing the leaf [i.e., leaf economic spectrum, @Wright2004a; @Westoby2013].
Leaf mass per area (LMA or 1/SLA [specific leaf area]), determined by lamina thickness and leaf tissue density (LD), describes how much biomass is invested into given photosynthetic leaf area, which is a key trait in the leaf economic spectrum [@Wright2004a; @Poorter2009; @Onoda2017].
Generally, resource-acquisitive species (fast-growing species) tend to have low LMA values, showing high photosynthesis, high nutrients, and often have fast leaf turnover [@Wright2004a].
In contrast, resource-conservative species (slow-growing species) often have higher LMA values with the opposite patterns [@Garnier1994; @Wright2004a; @Reich2014; @Diaz2016].
The LMA is also frequently used in plant growth analysis [@Evans1972; @Poorter2014; @Falster2018] because relative growth rates (RGR) can be decomposed into the product of net assimilation rate (NAR), leaf mass ratio (LMR) and LMA (i.e., RGR = NAR $\times$ LMR $\times$ LMA^-1^).
Another appealing feature of LMA is that it is relatively easy to measure large numbers of species [i.e., 'soft' trait, @Diaz2004].
Therefore, LMA has been of interests to ecologists and widely used since the first report more than a century ago [@Hanson1917].
Actually, LMA has one of the best species coverage in leaf traits of the TRY plant data database (16,460 species) and has been also the most often requested leaf trait (2,977 out of 7,330 requests), followed by leaf by leaf nitrogen (N) contents per leaf dry mass [12,238 species, 1,938 requests, @Kattge2020 [TRY version 5, status October 1 2019]]

@@

In this study, we aimed to investigate the relationship between whole-leaf (including petiole) LMA and leaf disc LMA,
whether the relationship varied with leaf tissue density, leaf area and leaf thickness, and
whether the extent of ITV for each species match between whole-leaf and leaf disc based estimates.
We collected a total of `r val$all_sp` woody species from four biomes (tropics, subtropics, savanna, and warm-temperate) to cover the wide range of geography.
We evaluate the following hypotheses:
(1) whole-leaf LMA would be greater than leaf disc LMA because whole-leaf LMA includes petioles and midribs that have greater dry mass per unit area than laminas [@Niinemets2007],
(2) species with higher leaf tissue density, larger and/or thinner leaves would show larger variance and differences in LMA estimates because whole-leaf LMA should be driven more in those species by veins and petioles , whereas those effects are likely to be ignored in leaf disc samples, and
(3) the extent of ITV does not match well between whole-leaf and leaf disc based estimates because leaf disc samples have more sources of variation than whole-leaf samples.
In addition to leaf morphology, we also investigate the effects of leaf punch size on the relationship between whole-leaf and leaf disc LMA.

@@

There was a significant shift in elevation between whole-leaf and leaf disc LMA (Fig.1, SMA slope: `r val$sma_slope_all_mean` [95%CI: `r val$sma_slope_all_lwr`, `r val$sma_slope_all_upr`], SMA intercept: `r val$sma_int_all_mean` [95%CI: `r val$sma_int_all_lwr`, `r val$sma_int_all_upr`]), suggesting that the whole-leaf estimates of LMA were `r val$sma_per_all_mean`% [95%CI: `r val$sma_per_all_lwr`, `r val$sma_per_all_upr`] greater than disc-based estimates in the arithmetic scale.

@@

Leaf disc LMA explained interspecific variations in whole-leaf LMA (*R^2^* = `r val$r2_lma`, *n* = `r val$all_sp`) quite well, and the strength of their relationship was approximately same or slightly weaker than previously reported  [*R^2^* = 0.92, *n* = 409, for whole-leaves in @Kraft2008; *R^2^* = 0.92, *n* = 364 for leaf laminas in @Onoda2011].
Standardized major axis (SMA) regression analysis showed significantly greater intercepts when all the species were pooled (`r val$sma_per_all_mean`% differences).
Together, these findings suggest that disc-based estimates can be generally used as good proxies for their whole-leaf estimates with appropriate calibration (see discussion below).
However, leaf tissue density, leaf area, leaf thickness and leaf punch size also influenced the relationship between the whole-leaf LMA and leaf disc LMA (Fig. 2 and Appendix S4).

@@

The effect of leaf punch size was the most important source of the variation (Fig. 2, Appendix S4), and samples obtained with a small leaf punch showed greater variance in whole-leaf to leaf disc LMA ratio.
Small leaf disc size inflates trait varition within leaves because dry mass will be inflated when major veins are unintentionally included in the samples.
Species with lower leaf tissue density and thin leaves showed greater variance in whole-leaf to leaf disc LMA ratio especially when a small punch was used (Fig.2ac: small leaf punch, Fig. 2a: large leaf punch).
@Kraft2008 found high *R^2^* values (*R^2^* = 0.92) between whole-leaf and leaf disc LMA based on leaf discs of approximately 1.0-cm diameter for sapling leaves under closed canopy that are softer and thinner than adult leaves [@Kitajima2010; @Oktavia2020].
Altough a leaf punch of 1.0-cm diameter was used only in the Yakushima dataset in our study, the variance in the whole-leaf to leaf disc LMA ratio were small for the Yakushima dataset, regardless of leaf tissue density or leaf thickness (Fig. 2ac).
On the other hand, a previous study found that leaf disc LMA obtained with a very small leaf punch (0.3-cm diameter) and whole-leaf LMA had different relationships between shade tolerance [@Poorter2009a], suggesting that leaf disc LMA and whole-leaf LMA can be inherently different when small disc samples were used.
The relatively low *R^2^* values in the Yunnan dataset (Appendix S5) compared to the previous study and the Yakushima dataset may suggest not only the effect of distinct veins on variance in LMA [@John2017] or distinct veins with kite-type appearances [@Grubb1986] but also technical difficulty in measuring tiny weight of discs in lower LMA leaves.
We found that estimates of whole-leaf and leaf disc LMA diverged when total dry mass of leaf discs was small (Appendix S6), and thus species with lower leaf tissue density and thin leaves, which have smaller dry mass (Appendix S7), tended to show grater variation in whole leaf to leaf disc LMA ratio especially when we used a smaller leaf punch.

@@

When the objective is to use leaf disc LMA as an alternative for whole-leaf LMA, using an appropriate size of leaf punch (e.g., 1.0-cm in diameter) is important for obtaining stable estimates of leaf disc LMA that matches well with whole-leaf LMA.
Our results suggest that simple conversion between whole-leaf and leaf disc LMA is difficult for species obtained with a small leaf punch.
The effects of leaf tissue density and leaf area are predictable when we used a large leaf punch (Fig. 2ab: large leaf punch), suggesting that conversion between leaf disc and whole-leaf LMA should be possible using leaf area as a covariate (measuring leaf tissue density requires whole-leaf LMA itself).
The ordinary least-squares (OLS) regressions with leaf area as a covariate as following:
(1) log~10~LMA~w~ = `r val$yaku2_slope_disc` $\times$ log~10~LMA~d~ + `r val$yaku2_slope_area` $\times$ log~10~LA,
(2) log~10~LMA~d~ = `r val$yaku2_int_inv_log10` + `r val$yaku2_slope_inv_leaf` $\times$ log~10~LMA~w~  - 0.063  $\times$ log~10~LA (*R^2^* = `r val$yaku_r2_area`, *AIC* = `r val$yaku_aic_area`, *n* = `r val$yaku_sp`),
where LMA~w~, LMA~d~, and LA are the whole-leaf LMA, leaf disc LMA and leaf area, respectively.
Thej OLS regressions without leaf area as a covariate as following:
(3) log~10~LMA~w~ = `r val$yaku_int_log10` + `r val$yaku_slope_disc` $\times$ log~10~LMA~d~ or
(4) log~10~LMA~d~ = `r val$yaku_slope_inv_leaf` $\times$ log~10~LMA~w~
(*R^2^* = `r val$yaku_r2_simple`, *AIC* = `r val$yaku_aic_simple`, *n* = `r val$yaku_sp`).
Nonetheless, including leaf area only slightly improves the *R^2^* values for the Yakushima dataset where a large punch was used.
This is because the large discrepancy in the whole-leaf to leaf disc LMA for large leaves can be explained by scaling relationships between leaf mass and leaf area called "diminishing returns" [@Niklas2007] which refers to increases in leaf dry mass including petioles do not result in the proportional increase in leaf area including petioles [@Li2022].
This effect is generally strong for mature leaves [@Jiao2022].
The "diminishing returns" can be either described as the scaling slope of leaf area in the model 1 or the scaling slope of leaf-disc LMA and the intercept in the model 3.
When one needs to mix leaf disc LMA and whole-leaf LMA in a single analysis (e.g., meta-analysis) or quickly measure leaf disc LMA as a proxy for whole-leaf LMA, we recommend to convert LMA values using one of the models above.

@@

# Conclusions
